---
title: "Data Story"
description: |
  Investigation of world fuel consumption and carbon dioxide (CO<sub>2</sub>) emissions from 1990-2020 using a data set obtained from [Enerdata](https://yearbook.enerdata.net/). 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo = FALSE}
# Loading in the data set and packages
library(ggplot2)
library(gganimate)
library(gifski)
library(png)
library(av)
energy = read.csv("energy_data_set.csv", header = TRUE)
```

```{r, echo = FALSE}
# Cleaning the data set

# Changing column names
colnames(energy) = c("country", "Year", "Region", "co2_fuel_com", "avg_co2",
                     "co2_gdp", "tot_energy_prod", "tot_energy_con", "perc_renew",
                     "elec_demand", "oil_dom_con", "ref_oil_prod", "gas_prod",
                     "gas_dom_con", "con_gdp", "elec_prod", "elec_dom_con", 
                     "coal_dom_con", "perc_wind_sol", "crud_oil_prod", "coal_prod")

# Converting variables to appropriate class
# Making region a factor
energy$Region = factor(energy$Region, 
                       levels = c("Arab States", "South/Latin America", 
                                  "Asia & Pacific", "Europe", "North America",
                                  "Middle east", "Africa"))
# Making numeric vectors
energy$gas_prod = as.numeric(energy$gas_prod)
energy$coal_dom_con = as.numeric(energy$coal_dom_con)
energy$perc_wind_sol = as.numeric(energy$perc_wind_sol)
energy$coal_prod = as.numeric(energy$coal_prod)

# Considering only complete cases
energy = energy[complete.cases(energy), ]
```

As greenhouse gasses, like CO<sub>2</sub>, are emitted from processes such as fossil fuel combustion, heat is trapped in our atmosphere, contributing to climbing temperatures and more extreme weather events; in other words, climate change. Shifting to cleaner, renewable energy sources with smaller carbon footprints offers one avenue to reduce our emissions of greenhouse gasses. To look more closely at global CO<sub>2</sub> emissions and fuel consumption, I will be presenting data from [Enerdata's](https://yearbook.enerdata.net/) data set, "World energy data 1990-2020." Something to note before moving on is that, in this data set, the only country in the Arab states for which we have data is Algeria.


### How have average total global CO<sub>2</sub> emissions compared to average global CO<sub>2</sub> emissions from fuel combustion (1990-2020)?

```{r}
year = c(1990:2020)

avg_tot_co2_global = NULL
for(i in 1:nrow(energy)){
  for(y in 1:31){
    avg_tot_co2_global[y] = mean(energy$avg_co2[energy$Year == year[y]])
  }
}

avg_co2_combust_global = NULL
for(i in 1:nrow(energy)){
  for(y in 1:31){
    avg_co2_combust_global[y] = mean(energy$co2_fuel_com[energy$Year == year[y]])
  }
}

# Plotting
 ggplot() +
  geom_point(aes(x = year, y = avg_tot_co2_global, size = avg_co2_combust_global), color = "#F8500C", alpha = 0.75) +
  labs(x = "Year",
       y = "Average total global CO2 emissions (ton/Mtoe)",
       size = "Average global CO2 emissions \n from fuel combustion (Mt)",
       caption = "Mtoe = million tons of oil equivalent") +
   ggtitle("Average Total Global CO2 Emissions and Average CO2 \n Emissions From Fuel Combustion (1990-2020)") +
   theme(plot.caption = element_text(color = "#6B6B6B", face = "italic"))
```

From 1990 to 2020, we observe a decrease in average total global CO<sub>2</sub> emissions from roughly 2.27 ton/Mtoe to roughly 2.13 ton/Mtoe. However, average global CO<sub>2</sub> emissions from fuel combustion have increased since 1990: over the 31 years from 1990 to 2020, there was an increase from 475 Mt to 722 Mt of CO<sub>2</sub> emitted from fuel combustion.

```{r}
library(gganimate)

anim = ggplot(energy, aes(x = co2_fuel_com, y = after_stat(density), fill = Region)) +
  geom_histogram(bins = 25, aes(fill = after_stat(count))) +
  labs(y = "Density",
       x = "CO2 emissions from fuel combustion (Mt)",
       subtitle = "Year: {frame_time}",
       caption = "Mt = million tons") +
  ggtitle("CO2 Emissions From Fuel Combustion by Region") +
  scale_fill_continuous(low = "orange",
                       high = "red",
                       name = "Number of \n countries") +
  facet_wrap(~ Region) +
  theme(plot.caption = element_text(color = "#6B6B6B", face = "italic")) +
  transition_time(energy$Year) +
  ease_aes("linear")

animate(anim, nframes = 300)
```

When looking at the how the distribution of CO<sub>2</sub> emissions from fuel combustion changes from 1990-2020 for each region in the histograms above, it is immediately obvious that the most substantial change occurs in Asia & Pacific. The bar that inches towards 10,000 Mt of CO<sub>2</sub> emissions from fuel combustion represents China which increased their emissions from 2,255.4 Mt in 1990 to 9,716.772 Mt in 2020.

<br>

```{r}
library(flextable)

region.name <- c("Africa", "Arab States", "Asia & Pacific",
                                 "Europe", "Middle east", "North America",
                                 "South/Latin America")
avg_co2_com_1990 = NULL
for(i in 1:nrow(energy)){
  for(y in 1:7){
    if(energy$Region[i] == region.name[y]){
      avg_co2_com_1990[y] = mean(energy$co2_fuel_com[energy$Region == region.name[y] & energy$Year == 1990])
    }
  }
}

avg_co2_com_2020 = NULL
for(i in 1:nrow(energy)){
  for(y in 1:7){
    if(energy$Region[i] == region.name[y]){
      avg_co2_com_2020[y] = mean(energy$co2_fuel_com[energy$Region == region.name[y] & energy$Year == 2020])
    }
  }
}



tab.df <- data.frame(Region = region.name,
                    avg_co2_com_1990 = round(avg_co2_com_1990, 1),
                    avg_co2_com_2020 = round(avg_co2_com_2020, 1),
                    Change = round(avg_co2_com_2020 - avg_co2_com_1990, 1))
table <- flextable(tab.df)
table <- add_header_row(x = table, values = c("", "Average CO2 Emissions from Fuel
                                              Combustion (Mt)", ""),
                 colwidths = c(1, 2, 1))
table <- align(table, i = 1, part = "header", align = "center")
table <- labelizor(table, part = "header",
                 labels = c("avg_co2_com_1990" = "1990",
                            "avg_co2_com_2020" = "2020",
                            "Change" = "Change"))
table <- autofit(table)
table <- footnote(table, i = 1, j = 2,
                  value = as_paragraph("Mt = million tons"),
                  ref_symbols = "a",
                  part = "header")
table
```


When we take a closer look at average CO<sub>2</sub> emissions from fuel combustion by region, we see that Asia & Pacific had the most dramatic increase in average emissions from 453.2 Mt in 1990 to 1,398.1 Mt in 2020. We also see that North America has the highest average CO<sub>2</sub> emissions from fuel combustion by far, but it has decreased from 2,648 Mt in 1990 to 2,460.1 Mt in 2020. Europe also managed to slightly lower average CO<sub>2</sub> emissions from fuel combustion from 461 Mt in 1990 to 322 Mt in 2020. The rest of the world had smaller average CO<sub>2</sub> emissions from fuel combustion, all under 250 Mt, but still increased their emissions from 1990 to 2020.


### Fuel Consumption in Different Regions of the World

Use the drop down menu below to view the domestic consumption of different fuel types around the world from 1990 to 2020.


<iframe height="760" width="100%" frameborder="no" src="https://rconnect.bucknell.edu/content/6a83d508-3719-4733-a265-529618cbcc08"> </iframe>


### How Does the Percentage of Electricity Produced by Renewable Resources Impact Overall CO<sub>2</sub> Emissions?

```{r, fig.height=6, fig.width=8}
library(ggdensity)
ggplot(energy, aes(x = perc_renew, y = avg_co2)) +
    geom_hdr() +
  geom_point(size = 0.2, color = "#F8500C") +
  geom_smooth(method = "lm", linewidth = 0.3, color = "red") +
  facet_wrap(~ Region) +
  geom_hdr_rug(aes(fill = after_stat(probs)), alpha = 1, sides = "lb") +
  scale_fill_brewer(palette = "Reds") +
  guides(alpha = "none") +
  labs(y = "Total CO2 emission (tons/Mtoe)",
       x = "Percent of total electricity production that is renewable",
       caption = "Mtoe = million tons of oil equivalent") +
  ggtitle("Total CO2 Emissions vs Percentage of Electiricity Production That is \n From Renewable Resources (1990-2020)") +
  theme(plot.caption = element_text(color = "#6B6B6B", face = "italic"))

```

As expected, on average, a higher percentage of the total electricity produced in a region correlates to lower total CO<sub>2</sub> emissions as per the negative correlations suggested by the fitted LOESS lines.

### Summary


It is crucial for the emission of greenhouse gasses, like CO<sub>2</sub>, to be reduced to slow the acceleration of climate change by anthropogenic activities such as the burning of fossil fuels like oil, natural gas, and coal. While climate change is a challenge that must be tackled by nations all over the world, the data from Enerdata suggest that, on average as of 2020, North America is the largest consumer of oil and natural gas. Notably, China has substantially increased their domestic coal consumption from 1990 to 2020 making the Asia & Pacific the largest consumers of domestic coal. The radar plot shown below shows more recent (2015-2020) relative average fossil fuel consumption between different regions.



```{r}
library(ggradar)

region.name = unique(energy$Region)


 oil_con = 1:7
 for(i in 1:nrow(recent_data)){
   for(y in 1:7){
     if(recent_data$Region[i] == region.name[y]){
       oil_con[y] = mean(recent_data$oil_dom_con[recent_data$Region == region.name[y]])
     }
   }
 }
 
 oil_con = oil_con/max(oil_con)

 
 coal_con = 1:7
 for(i in 1:nrow(recent_data)){
   for(y in 1:7){
     if(recent_data$Region[i] == region.name[y]){
      coal_con[y] = mean(recent_data$coal_dom_con[recent_data$Region == region.name[y]])
    }
  }
 }
 
 coal_con = coal_con/max(coal_con)

gas_con = 1:7
for(i in 1:nrow(recent_data)){
  for(y in 1:7){
    if(recent_data$Region[i] == region.name[y]){
      gas_con[y] = mean(recent_data$gas_dom_con[recent_data$Region == region.name[y]])
    }
  }
}



gas_con = gas_con/max(gas_con)

radar.con = data.frame(Region = region.name,
                        oil_con = oil_con,
                        coal_con = coal_con,
                        gas_con = gas_con)

 ggradar(radar.con, group.line.width = .5, group.point.size = 1,
         grid.label.size = 0, axis.label.size = 3, axis.labels = c("Domestic Oil Consumption",
                         "Domestic
 Coal
 Consumption",
                         "Domestic
                         Natural Gas
                         Consumption")) +
   theme(legend.text = element_text(size = 9),
         plot.title = element_text(size = 13)) +
   ggtitle("Comparison of Relative Average Fuel 
           Consumption (2015-2020)")
```



As of 2015-2020, South/Latin America and North America are world leaders in the production of cleaner electricity, each with greater than an average 40% of electricity being generated from renewable resources. Europe and Asia & Pacific follow with an average of 30.8% and 20.5% of electricity that is renewable, respectively.



```{r}

recent_data = energy[energy$Year >= 2015 & energy$Year <= 2020, ]

recent_renew = NULL
for(i in 1:nrow(recent_data)){
  for(y in 1:7){
    if(recent_data$Region[i] == region.name[y]){
      recent_renew[y] = mean(recent_data$perc_renew[recent_data$Region == region.name[y]]) 
    }
  }
}


sd_recent_renew = NULL
for(i in 1:nrow(recent_data)){
  for(y in 1:7){
    if(recent_data$Region[i] == region.name[y]){
      sd_recent_renew[y] = sd(recent_data$perc_renew[recent_data$Region == region.name[y]]) 
    }
  }
}

region.name = unique(energy$Region)

ggplot() +
  geom_bar(aes(x = region.name, y = recent_renew), stat = "identity",
           fill = "#F8500C") +
  geom_errorbar(aes(x = region.name, ymin = recent_renew - sd_recent_renew,
                    ymax = recent_renew + sd_recent_renew), width = 0.2) +
  labs(x = "Region",
       y = "Average Percent of Renewable Electricity",
       caption = "Error bars represent the standard deviation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size = 9),
        plot.caption = element_text(color = "#6B6B6B", face = "italic")) +
  ggtitle("Average Percent of Renewable Electricity in Each Region (2015-2020)")
```




